---
title: "Replication of Study 1 by Smiley & Fisher (2022, Psychological Science)"
author: 
  name: "Sarah H. Wu (sarwu@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format: 
  html: 
    toc: true 
    toc_depth: 3
editor: 
  markdown: 
    wrap: 72
---

<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

## Introduction

Despite the idiosyncracies of individual technological developments,
there is a fundamental human panic about new technologies. In the 1800s,
novels were accused of corrupting the minds of youth and housewives, in
the 1940s "television addiction" was coined, and in current discourse
smartphones and social media are the supposed culprits behind
adolescents' declining mental health. (See [Pessimists
Archive](https://pessimistsarchive.org/) for more examples.) Because my
research investigates people's perceptions of digital media, I found
Smiley & Fisher (2022) relevant because it probes a potential mechanism
underlying cycles of tech panic: status quo bias, which has been
documented across domains and developmental stages. In their Study 1,
Smiley and Fisher (2022) experimentally manipulated participants' age at
the time of an obscure technology's invention (15 years before vs. after
the invention of the technology) and measured participants'
self-reported perceptions of the technology's societal impact. Smiley
and Fisher (2022) found that participants in the born-after condition
evaluated the technology more positively than those in the born-before
condition.

The stimuli for this experiment are descriptions of an obscure
technology called aerogel, "a synthetic porous ultralight material
derived from a gel, in which the liquid component for the gel has been
replaced with a gas without significant collapse of the gel structure."
The authors selected this technology because it was possible to
manipulate its perceived age without too many participants knowing its
actual date of invention. In the born-before condition, participants
were told that aerogel was invented 15 years after they had been born,
with the exact year dynamically populated based on the participant's
birth year. (For example, a participant born in 2000 would read that
aerogel was invented in 2015.) In the born-after condition, participants
were told that aerogel had been invented 15 years prior to their birth.
(So a participant born in 2000 would see that aerogel was invented in
1985.) After reading about aerogel, participants were asked, "How would
you describe the impact of aerogel on society?" and reported their
impression on a 7-point Likert scale from -3 (very negative) to 3 (very
positive). Then, they answered a manipulation check question and a
previous knowledge question.

The experiment seems relatively straightforward to replicate. However,
as with any online study, it is possible that participants might look up
details about the stimuli during the study, which could render the
manipulation ineffective. Because aerogel is a real invention, people
could easily find its actual date of invention.

Original paper
[here.](https://github.com/psych251/smiley2022/blob/main/original_paper/smiley-fisher-2022-the-golden-age-is-behind-us-how-the-status-quo-impacts-the-evaluation-of-technology.pdf)

## Methods

### Power Analysis

The original effect size from Smiley and Fisher (2022) was *d* = .25. As
in the original paper, I aimed for 80% power with a medium effect size
of *d* = .3. G\*Power confirmed that this would take at least 176
participants in each condition.

### Planned Sample

The planned sample size is 400 participants, using a representative
sample based on US census data.

### Materials

The materials and procedure were followed precisely.

"To begin, participants provided demographic information: age, gender,
and level of education. Next, participants were told, “Aerogel (pictured
above) \[see Section S1 in the Supplemental Material at
https://osf.io/bxjnr\] is a synthetic porous ultralight material derived
from a gel, in which the liquid component for the gel has been replaced
with a gas without significant collapse of the gel structure.” This
technology was selected because, though real, it is largely unknown,
making it possible to manipulate its perceived age."

### Procedure

From the original paper: "Participants were randomly assigned to the
born-before or born-after condition. In the born-before condition,
participants were next told that aerogel was invented 15 years after
they had been born. We dynamically presented the invention date based on
each participant’s age. For example, a participant born in 1980 would
see, “It was invented in 1995 and is now widely used.” In the born-after
condition, participants were told that aerogel had been invented 15
years before they were born. For example, a participant born in 1980
would see, “It was invented in 1965 and is now widely used.” After
viewing this information, all participants were asked, “How would you
describe the impact of aerogel on society?” and responded on a 7-point
Likert scale ranging from −3 (very negative) to 3 (very positive).
Finally, as a manipulation check, participants were asked, “Was aerogel
invented before or after the year you were born?” (before/after), and a
previous knowledge question asked, “Did you know about aerogel prior to
completing this survey?” (yes/no)."

### Analysis Plan

The key analysis of interest is an unpaired t-test comparing evaluations
of aerogel in the born-before and born after conditions. Like the
original paper, we will compare the size of the effect with and without
exclusions based on the manipulation check and prior knowledge about
aerogel.

In line with the original paper, we will also conduct a zero-order
correlation to test a competing hypothesis: that people simply prefer
older technology. By this account, older people would tend to have more
positive evaluations because aerogel was described as especially old
(either 15 years before or after their birth).

### Differences from Original Study

The main difference between this replication study and the original
study is that the original study used an MTurk sample, whereas we used a
Prolific sample. Prolific is generally recognized as having
higher-quality participants (Douglas et al., 2023), but these
differences are probably negligible given the ease of the task.

### Methods Addendum (Post Data Collection)

#### Actual Sample

My actual sample size was 394 after excluding participants who failed
the initial attention check. Like Smiley and Fisher (2022), I also
tracked who failed the manipulation check (N = 37) and who had prior
knowledge of aerogel (N = 69), analyzing the data both with and without
these exclusions.

#### Differences from pre-data collection methods plan

None.

## Results

### Pilot B: Data preparation

Data preparation following the analysis plan.

```{r, echo=FALSE}
### Data Preparation

#### Load Relevant Libraries and Functions
if(!require("psych")) {install.packages("psych"); require("psych")}
if(!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if(!require("effsize")) {install.packages("effsize"); require("effsize")}
if(!require("stats")) {install.packages("stats"); require("stats")}
if(!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
library(qualtRics)
library(VennDiagram)
library(grid)
library(ggvenn)
library(rstatix)
library(interactions)
library(patchwork)
library(papaja)
library(gtsummary)
```

```{r}
#### Impgrid#### Import data
pilot_b <- read_csv("../data/pilot_b.csv") %>% 
  drop_na(id) %>%  # Only include Prolific participants, not Pilot A participants
  filter(DistributionChannel=="anonymous")


#### Data exclusion / filtering
pilot_b_recoded <- pilot_b %>% 
  rename(Condition=FL_9_DO,
         Eval = attitude)

### DATA WRANGLING & CLEANING ###

## Participant Info
pilot_b_recoded$gender <- factor(pilot_b_recoded$gender, levels = c("1", "2"), labels=c("Male","Female"))


pilot_b_recoded$age <- as.numeric(pilot_b_recoded$age)
pilot_b_recoded$Eval <- as.numeric(pilot_b_recoded$Eval)

### END DATA WRANGLING & CLEANING ###

## Demographic information
table(pilot_b_recoded$gender)
mean(pilot_b_recoded$age)
sd(pilot_b_recoded$age)

```

### Pilot B: Confirmatory analysis

```{r}
### Analysis WITHOUT exclusions ###
## Descriptive statistics
s1.cond.difs <- pilot_b_recoded %>% 
  group_by(Condition) %>% 
  summarize(EVAL = mean(Eval), EVAL_SD = sd(Eval))
s1.cond.difs

## T-test and Cohen's d
## Condition refers to born after/before invention (experimentally manipulated)
t.test(Eval ~ Condition, var.equal=TRUE, data=pilot_b_recoded)
effsize::cohen.d(Eval ~ Condition, data = pilot_b_recoded)

# Pearson's correlation
cor.test(pilot_b_recoded$age, pilot_b_recoded$Eval)

### END Analysis WITHOUT exclusions

### Analysis WITH exclusions ###
## Exclusion
table(pilot_b_recoded$attn_check_1) # 
table(pilot_b_recoded$attn_check_2)
table(pilot_b_recoded$prior_knowledge) # 
# Exclude participants who failed manipulation check or knew what aerogel was
s1.df.f <- pilot_b_recoded %>% 
  filter(attn_check_1 == "3" & attn_check_2 == "2") %>% 
  filter(prior_knowledge != "1")

## Descriptive statistics
s1.cond.difs <- s1.df.f %>% 
  group_by(Condition) %>% 
  summarise(EVAL = mean(Eval), EVAL_SD = sd(Eval))

s1.cond.difs
```

```{r}

## T-test and Cohen's d 
t.test(Eval ~ Condition, var.equal=TRUE, data=s1.df.f)
effsize::cohen.d(Eval ~ Condition, data = s1.df.f)

## Age and Evaluation correlation
cor.test(s1.df.f$age, s1.df.f$Eval)

## Make Condition.Num numeric so it can be scaled in lm
s1.df.f$Condition.Num <- rep(NA, length(s1.df.f$Condition))
s1.df.f$Condition.Num[s1.df.f$Condition=="AerogelBornAfter"] <- 0
s1.df.f$Condition.Num[s1.df.f$Condition=="AerogelBornBefore"] <- 1

## Age and Condition multilevel model
s1.lm.age <- lm(scale(Eval) ~ scale(age) + scale(Condition.Num), data = s1.df.f)
summary(s1.lm.age)

```

### Full Dataset: Data Preparation

```{r, echo=FALSE}
full_sample <- fetch_survey(surveyID = "SV_ahlhnHc1n0Mfmqq",
                            label = TRUE,
                            convert=FALSE,
                            start_date = "2025-12-01 21:13:39")
```

```{r}
full_sample_conditions <- full_sample %>% 
  mutate(Condition = if_else(is.na(FL_9_DO_AerogelBornBefore), "Born-After", "Born-Before")) %>%  # Create condition column 
  filter(attn_check_1 == "50" & attn_check_2 == "Her mother") %>% # Filter out those who failed attention check and didn't get to finish
  filter(Finished==1) %>% 
  mutate(
    attitude_num = as.numeric(str_extract(attitude, "^[^a-zA-Z]+"))
  ) %>% 
  relocate(attitude_num, .after = attitude)
  
```

### Demographics

```{r}

demographics_table <- full_sample_conditions %>%
  tbl_summary(
    include = c(age, gender),            # Variables to include
    label = list(
      age ~ "Age (Years)",               # Custom Label for Age
      gender ~ "Gender Identity"         # Custom Label for Gender
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})", # Default is Median (IQR); change to Mean (SD)
      all_categorical() ~ "{n} ({p}%)"    # Count and Percentage
    ),
    digits = all_continuous() ~ 1        # Round continuous stats to 1 decimal place
  ) %>%
  add_n() %>%                            # Add column with total N
  gtsummary::modify_header(label = "**Variable**")  # Rename the stub header

# 3. Print the table
demographics_table
```

```{r}
### Analysis WITHOUT exclusions ###
## Descriptive statistics
condition_diffs_no_exclusion <- full_sample_conditions %>% 
  group_by(Condition) %>% 
  get_summary_stats(attitude_num, type="mean_ci")
condition_diffs_no_exclusion

full_sample_conditions %>% 
  group_by(Condition) %>% 
  get_summary_stats(attitude_num, type="mean_sd")

## T-test and Cohen's d
## Condition refers to born after/before invention (experimentally manipulated)
without_exclusions <- t.test(attitude_num ~ Condition, var.equal=TRUE, data=full_sample_conditions)
effsize::cohen.d(attitude_num ~ Condition, data = full_sample_conditions)
without_exclusions_apa <- apa_print(without_exclusions)

# Pearson's correlation
cor.test(full_sample_conditions$age, full_sample_conditions$attitude_num)

# Distribution of ages
ggplot(full_sample_conditions, aes(x=age)) +
  geom_histogram(binwidth=5, color="white") +
  ylim(c(0,60)) +
  labs(x="Age") +
  jtools::theme_apa()
```

Participants in the born-before condition evaluated aerogel no more
favorably than participants in the born-after condition,
`r without_exclusions_apa$full_result`

```{r}
# Exclude participants who failed manipulation check or knew what aerogel was
full_sample_exclusions <- full_sample_conditions %>% 
  filter(manip_check == Condition) %>% # Their answer should match the condition to which they were assigned
  filter(prior_knowledge != "Yes") # They should not know what aerogel is

## Descriptive statistics
condition_diffs_exclusions <- full_sample_exclusions %>% 
  group_by(Condition) %>% 
  get_summary_stats(attitude_num, type="mean_ci")

condition_diffs_exclusions

## T-test and Cohen's d 
with_exclusions <- t.test(attitude_num ~ Condition, var.equal=TRUE, data=full_sample_exclusions)
with_exclusions_apa <- papaja::apa_print(with_exclusions)
effsize::cohen.d(attitude_num ~ Condition, data = full_sample_exclusions)

## Age and Evaluation correlation
cor.test(full_sample_exclusions$age, full_sample_exclusions$attitude_num)
```

In line with Smiley and Fisher (2022), I also analyzed the data after
excluding participants who failed the manipulation check or had prior
knowledge of aerogel.

Still, participants in the born-before condition evaluated aerogel no
more favorably than those in the born-after condition,
`r with_exclusions_apa$full_result`

```{r}
# Visualizations for original study
original_data <- read_csv("../data/TechAgeS1.csv")



ggplot(original_data, aes(x=Age)) +
  geom_histogram(binwidth=5, color="white") +
  jtools::theme_apa()

# Run stats
stat.test <- original_data %>% 
  t_test(Eval ~ Condition) %>% 
  add_significance() %>% 
  add_xy_position(x = "Condition")

smiley2022_plot <- original_data %>%
  group_by(Condition) %>% 
  get_summary_stats(Eval, type="mean_ci") %>% 
  ggplot(aes(x=Condition, y=mean, color=Condition)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.2) +
  geom_text(aes(label = paste0("italic(N) == ", n), y = mean + ci + 0.3), 
            parse=TRUE,
            color = "black", 
            size = 4) +
  ggpubr::stat_pvalue_manual(stat.test, 
                     label = "p.signif", # Uses stars (*, **, ns)
                     y.position = 3.2,   # Adjust height
                     color = "black") +
  scale_y_continuous(limits = c(-3, 3.5), breaks = seq(-3, 3, 1)) +
  scale_color_manual(values=c("steelblue", "orange")) +
  labs(y="Evaluation of Aerogel",
       title= "Smiley & Fisher (2022)") +
  jtools::theme_apa() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12),
        plot.title = element_text(hjust=.5)) 

stat.test.exclusions <- original_data %>% 
  filter(Check=="Pass" & Knowledge==2) %>% 
  rstatix::t_test(Eval ~ Condition) %>% 
  add_significance() %>% 
  add_xy_position(x = "Condition")

original_data %>%
  filter(Check=="Pass" & Knowledge==2) %>% 
  group_by(Condition) %>% 
  get_summary_stats(Eval, type="mean_ci") %>% 
  ggplot(aes(x=Condition, y=mean, color=Condition)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.2) +
  geom_text(aes(label = paste0("italic(N) == ", n), y = mean + ci + 0.3), 
            parse=TRUE,
            color = "black", 
            size = 4) +
  ggpubr::stat_pvalue_manual(stat.test.exclusions, 
                     label = "p.signif", # Uses stars (*, **, ns)
                     y.position = 3.2,   # Adjust height
                     color = "black") +
  scale_y_continuous(limits = c(-3, 3.5), breaks = seq(-3, 3, 1)) +
  scale_color_manual(values=c("steelblue", "orange")) +
  labs(y="Evaluation of Aerogel") +
  jtools::theme_apa() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12))
```

```{r}
# Visualizations for my replication study
replication_plot <- ggplot(condition_diffs_no_exclusion, aes(x=Condition, y=mean, color=Condition)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.2) +
  geom_text(aes(label = paste0("italic(N) == ", n), y = mean + ci + 0.3), 
            parse=TRUE,
            color = "black", 
            size = 4) +
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-3, 3, 1)) +
  scale_color_manual(values=c("steelblue", "orange")) +
  labs(y="Evaluation of Aerogel",
       title="Replication Attempt") +
  jtools::theme_apa() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12),
        plot.title = element_text(hjust=.5)) 

ggplot(condition_diffs_exclusions, aes(x=Condition, y=mean, color=Condition)) +
  geom_point(size=3) +
  geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.2) +
  geom_text(aes(label = paste0("italic(N) == ", n), y = mean + ci + 0.3), 
            parse=TRUE,
            color = "black", 
            size = 4) +
  scale_y_continuous(limits = c(-3, 3), breaks = seq(-3, 3, 1)) +
  scale_color_manual(values=c("steelblue", "orange")) +
  labs(y="Evaluation of Aerogel") +
  jtools::theme_apa() +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 12)) 



## Make Condition.Num numeric so it can be scaled in lm
full_sample_exclusions$Condition.Num <- rep(NA, length(full_sample_exclusions$Condition))
full_sample_exclusions$Condition.Num[full_sample_exclusions$Condition=="Born-After"] <- 0
full_sample_exclusions$Condition.Num[full_sample_exclusions$Condition=="Born-Before"] <- 1

## Age and Condition multilevel model
full_sample_exclusions.lm.age <- lm(scale(attitude_num) ~ scale(age) + scale(Condition.Num), data = full_sample_exclusions)
summary(full_sample_exclusions.lm.age)
```

Neither condition nor age predicted attitudes toward aerogel.

#### Side-by-side plots

```{r}
smiley2022_plot + replication_plot
```

### Exploratory analyses

What are the characteristics of those who knew aerogel vs. those who did
not?

```{r}
# Education
full_sample_conditions %>% 
  group_by(education) %>% 
  count(prior_knowledge) %>% 
  mutate(prop=n/sum(n))

# Age
full_sample_conditions %>% 
  group_by(prior_knowledge) %>% 
  get_summary_stats(age, type="mean_sd")

full_sample_conditions %>% 
  group_by(prior_knowledge) %>% 
  get_summary_stats(`Duration (in seconds)`, type="mean_sd")
```

```{r}
# Examine potential overlap in the people who failed the manipulation check and those who knew aerogel
failed_manip <- full_sample_conditions %>% 
    filter(manip_check != Condition) 

knew_aerogel <- full_sample_conditions %>% 
  filter(prior_knowledge =="Yes")


ggvenn(list(
    "Failed Manipulation Check" = failed_manip$ResponseId, 
    "Knew About Aerogel" = knew_aerogel$ResponseId),
    set_name_size = 4,
    fill_color = c("darkred", "turquoise"))
```

```{r}
# Examine potential interaction between condition and age
ggplot(original_data, aes(x = Age, y = Eval, color = Condition)) +
  geom_point(alpha = 0.3) + # Raw data points
  geom_smooth(method = "lm", se = TRUE) + # Trend lines
  scale_color_manual(values=c("steelblue", "orange")) +
  labs(x="Age",
       y="Evaluation of Aerogel") +
  jtools::theme_apa()

# Test the interaction formally
model_interaction <- lm(Eval ~ Condition * Age, data = original_data)

# Check the summary for the "Condition:Age" term
summary(model_interaction)

ggplot(full_sample_conditions, aes(x = age, y = attitude_num, color = Condition)) +
  geom_point(alpha = 0.3) + # Raw data points
  geom_smooth(method = "lm", se = TRUE) + # Trend lines
  scale_color_manual(values=c("steelblue", "orange")) +
  labs(x ="Age",
       y="Evaluation of Aerogel") +
  jtools::theme_apa() 

model_interaction_replication <- lm(attitude_num ~ Condition * age, data = full_sample_conditions)

# Check the summary for the "Condition:Age" term
summary(model_interaction_replication)
```

## Discussion

### Summary of Replication Attempt

I failed to replicate the main finding from the original Study 1 from
Smiley & Fisher (2022), which was that participants in the born-after
condition had more favorable attitudes toward aerogel than participants
in the born-before condition. Smiley and Fisher (2022) documented this
effect both when they included participants who failed the manipulation
check and had prior knowledge of aerogel AND when they excluded those
participants. In my case, I did not find the effect in either case.

In my study, I also had to exclude a greater number of participants
because 38 participants failed the manipulation check and an additional
60 participants knew about aerogel before the study (vs. only 33
participants in the original study).

For those participants who knew about aerogel, they may not have
believed the information I presented to them about its year of
invention.

### Commentary

The fact that I failed to replicate the effect could have been due to a
few reasons: (1) the effect does not exist, (2) my manipulation of year
of aerogel invention was not as obvious as the original authors', (3) my
study was underpowered to detect an effect. My final sample size was
297, which is substantially smaller than the planned sample size.

On point (3), I noticed that the original authors were also underpowered
in their post-exclusion analysis of the effect of year of invention.
They determined they needed 176 participants per condition (352 total)
to detect a medium effect with 80% power, but after their exclusions
they had a total sample size of 328.

On a broader note, this (non) replication has reinforced the notion of
"playing 20 questions with nature" (Almaatouq et al., 2024). When an
effect cannot be replicated, it's difficult to tell if it's because the
replication involved a slightly different population/paradigm or because
the effect doesn't actually exist. Are we all just touching different
parts of the elephant? (And what is enough to constitute "different" in
the context of any experiment?)
